---
title: "Logistic HW 1"
output:
  html_document:
    df_print: paged
---

```{r, echo=F, message=F, warning=F}
library(MASS)
library(visreg)
library(brglm)
library(car)
library(mgcv)
library(haven)
library(tidyverse)
library(ggplot2)
library(DescTools)

file.dir <- "C:/Users/Jackson Cabell/Documents/Homework/Logistic/HW1/"
input.file1 <- "insurance_t.sas7bdat"

insurance <- read_sas(paste(file.dir, input.file1,sep = ""))
```

## Questions
```
Variables with less than 10 distinct levels are categorical? Ex: cashbk

Is NSF binary or ordinal?

ANOVA with 1-var model against null same as pearson chi sq? ANOVA() fcn with continous predictors? AKA are chi-sq tests only between two categorical variables.

Calculating odds ratio: still doing 1 variable comparison at a time?

Box-Tidwell func in R does vars one at a time regularly?
```


## Identify class of variables
```{r, message=F, warning=F}
#Continuous vars
cvars <- c(
  "ACCTAGE", "DDABAL", "DEP", "DEPAMT", "CHECKS", "NSFAMT", "PHONE", "TELLER", "SAVBAL", "ATMAMT", "POS", "POSAMT", "CDBAL", "IRABAL", "LOCBAL", "INVBAL", "ILSBAL", "MMBAL", "MTGBAL", "CCBAL", "INCOME", "LORES", "HMVAL", "AGE", "CRSCORE"
)

#Binary vars
bvars <- c(
  "DDA", "DIRDEP", "SAV", "ATM", "CD", "IRA", "LOC", "INV", "ILS", "MM", "MTG", "CC", "SDB", "HMOWN", "MOVED", "INAREA"
)

#Nominal vars
nvars <- c(
  "BRANCH", "RES"
)

#Ordinal vars cashbk mmcred ccpurc, nsf
ovars <- c(
  "CASHBK", "MMCRED", "CCPURC", "NSF"
)

#
vars <- data.frame("VAR" = c(bvars, nvars, cvars, ovars), "BINARY" = numeric(47), "NOMINAL" = numeric(47), "CONTINUOUS" = numeric(47), "ORDINAL" = numeric(47), "TYPE"= numeric(47))

vars <- vars %>% mutate(
  BINARY = ifelse(VAR %in% bvars, 1, 0),
  NOMINAL = ifelse(VAR %in% nvars, 1, 0),
  CONTINUOUS = ifelse(VAR %in% cvars, 1, 0),
  ORDINAL = ifelse(VAR %in% ovars, 1, 0)
) %>% mutate(
  TYPE = ifelse(BINARY==1, "BINARY", ifelse(NOMINAL==1, "NOMINAL", ifelse(CONTINUOUS==1, "CONTINUOUS", ifelse(ORDINAL==1, "ORDINAL", "ERROR"))))
)
```

## Test individual significance of each var: only using ANOVA
```{r, message=F, warning=F}
#Function to test significance of each var
signif <- function(x) {

VAR <- x[1]
CONTINUOUS <- x[4]

#Omit rows with missing data points for var tested
insurance_comp <- insurance[!is.na(eval(parse(text = paste("insurance$", VAR, sep="")))),]
  
nullmod <- glm(INS ~ 1, data = insurance_comp, family = binomial(link = "logit"))


#If not a continuous variable, need to call factor() function
if (CONTINUOUS == 0) {
  logit.model <-  glm(as.formula(paste("INS ~ factor(", VAR,")", sep="")), data = insurance_comp, family = binomial(link = "logit"))
} else {
  logit.model <-  glm(as.formula(paste("INS ~", VAR, sep="")), data = insurance_comp, family = binomial(link = "logit"))
}

#Do ANOVA
return(anova(logit.model, nullmod, test = 'LRT')[2,5])

#Alternatively, can do chisq.test(table(insurance_comp$INS, insurance_comp$VAR))

}

vars$SIGNIFICANCE <- apply(vars, 1, signif)

#Only vars with significance greater than 0.002
sigvars <- vars %>%
  filter(SIGNIFICANCE < .002) %>%
  arrange(SIGNIFICANCE) %>%
  select(VAR, TYPE, SIGNIFICANCE)

#write_csv(sigvars, paste(file.dir,"sigvars.csv",sep = ""))

#All vars ranked by significance
allvars <- vars %>%
  arrange(SIGNIFICANCE) %>%
  select(VAR, TYPE, SIGNIFICANCE)

write_csv(allvars, paste(file.dir,"allvars.csv",sep = ""))
allvars
```


## Test individual significance of each var: using ANOVA, Pearsons and MH Chi-Square Tests
```{r, message=F, warning=F}
#Function to test significance of each var
signif2 <- function(x) {

VAR <- x[1]
BINARY <- x[2]
NOMINAL <- x[3]
CONTINUOUS <- x[4]
ORDINAL <- x[5]

#Omit rows with missing data points for var tested
insurance_comp <- insurance[!is.na(eval(parse(text = paste("insurance$", VAR, sep="")))),]


#If not a continuous variable, need to call factor() function
if (CONTINUOUS == 1) {
  nullmod <- glm(INS ~ 1, data = insurance_comp, family = binomial(link = "logit"))
  logit.model <-  glm(as.formula(paste("INS ~", VAR, sep="")), data = insurance_comp, family = binomial(link = "logit"))
  return(anova(logit.model, nullmod, test = 'LRT')[2,5])
  
} else if(NOMINAL == 1) {
  return(chisq.test(table(insurance_comp$INS, eval(parse(text = paste("factor(insurance_comp$", VAR,")", sep="")))))$p.val)
} else {
  return(as.numeric(MHChisqTest(table(insurance_comp$INS, eval(parse(text = paste("factor(insurance_comp$", VAR,")", sep="")))))$p.val))
}


}

#Testing
#VAR <- "CD"
#insurance_comp <- insurance[!is.na(eval(parse(text = paste("insurance$", VAR, sep="")))),]
#chisq.test(table(insurance_comp$INS, eval(parse(text = paste("factor(insurance_comp$", VAR,")", sep="")))))$p.val #nominal
#as.numeric(MHChisqTest(table(insurance_comp$INS, eval(parse(text = paste("factor(insurance_comp$", VAR,")", sep="")))))$p.val) #ordinal

#nullmod <- glm(INS ~ 1, data = insurance_comp, family = binomial(link = "logit"))
  #logit.model <-  glm(as.formula(paste("INS ~", VAR, sep="")), data = insurance_comp, family = binomial(link = "logit"))
 # anova(logit.model, nullmod, test = 'LRT')

#Apply the function
vars$SIGNIFICANCE2 <- apply(vars, 1, signif2)

#Only vars with significance greater than 0.002
sigvars2 <- vars %>%
  filter(SIGNIFICANCE2 < .002) %>%
  arrange(SIGNIFICANCE2) %>%
  select(VAR, TYPE, SIGNIFICANCE2)

write_csv(sigvars2, paste(file.dir,"sigvars2.csv",sep = ""))

#All vars ranked by significance
allvars2 <- vars %>%
  arrange(SIGNIFICANCE2) %>%
  select(VAR, TYPE, SIGNIFICANCE2)

write_csv(allvars2, paste(file.dir,"allvars.csv",sep = ""))
allvars2
```


## Calculate individual odds ratios of binary vars
```{r, message=F, warning=F}
#Function to calculate odds ratios of each binary var
odds_ratio <- function(x) {

VAR <- x[1]

#Omit rows with missing data points for var tested
insurance_comp <- insurance[!is.na(eval(parse(text = paste("insurance$", VAR, sep="")))),]

#Fit model with binary variable
logit.model <-  glm(as.formula(paste("INS ~ factor(", VAR,")", sep="")), data = insurance_comp, family = binomial(link = "logit"))


#Calculate odds ratio
exp(
  cbind(coef(logit.model)[2], confint(logit.model))
)[2,]


}

binary_vars <- vars[vars$TYPE=="BINARY",]
binary_vars$ODDS <- numeric(nrow(binary_vars))
binary_vars$ODDS_LB <- numeric(nrow(binary_vars))
binary_vars$ODDS_UB <- numeric(nrow(binary_vars))

for(i in 1:nrow(binary_vars)) {
  binary_vars$ODDS[i] <- apply(binary_vars[i,], 1, odds_ratio)[1]
  binary_vars$ODDS_LB[i] <- apply(binary_vars[i,], 1, odds_ratio)[2]
  binary_vars$ODDS_UB[i] <- apply(binary_vars[i,], 1, odds_ratio)[3]
  
}

binary_vars <- binary_vars %>%
  arrange(desc(ODDS)) %>%
  select(VAR, TYPE, ODDS, ODDS_LB, ODDS_UB)

write_csv(binary_vars, paste(file.dir,"binary_vars.csv",sep = ""))
binary_vars
```

Interpretation of highest magnitude odds ratio:

INV (Indicator for investment account)
- Customers with an investment account have 3.4720391 times the odds of purchasing the new insurance product than those without an investment account.

Interesting findings:
- Those with checking accounts are about 62.5% less likely to purchase the insurance product.
- Those who interacted with ATM are 40.7% less likely to purchase the product.
- Credit card owners are 78% more likely.
- Home ownership is not significant (doesn't have an effect)
- Those with local address are 42.5% less likely.

## Linearity assumptions for continuous vars
```{r, message=F, warning=F}
#Function to test linearity assumption for continuous vars
gam_pval <- function(x) {

VAR <- x[1]

#Omit rows with missing data points for var tested
insurance_comp <- insurance[!is.na(eval(parse(text = paste("insurance$", VAR, sep="")))),]

#Fit model with cont variable
fit.gam <-  gam(as.formula(paste("INS ~ s(", VAR,")", sep="")), data = insurance_comp, family = binomial(link = "logit"), method = 'REML')

#Calculate pvalue
summary(fit.gam)$s.pv



}

#Filter variables into only continuous
cont_vars <- vars[vars$TYPE=="CONTINUOUS",]

#Add p-value to data
cont_vars$LIN_ASSUMP_PVAL <- apply(cont_vars, 1, gam_pval)

#Check if assumption is met and add it as variable
cont_vars <- cont_vars %>%
  select(VAR, TYPE, LIN_ASSUMP_PVAL) %>%
  arrange(desc(LIN_ASSUMP_PVAL)) %>%
  mutate(ASSUMPTION_MET = ifelse(LIN_ASSUMP_PVAL > .002, "YES", "NO"))

write_csv(cont_vars, paste(file.dir,"cont_vars.csv",sep = ""))
cont_vars

```


## Linearity assumptions for continuous vars
```{r, message=F, warning=F}
#Function to calc missing totals
missing_tot <- function(x) {

VAR <- x[1]

#Omit rows with missing data points for var tested
insurance_missing <- insurance[is.na(eval(parse(text = paste("insurance$", VAR, sep="")))),]

#Calculate number missing
nrow(insurance_missing)

}

vars$MISSING <- apply(vars, 1, missing_tot)

#All vars ranked by missing
vars_missing <- vars %>%
  arrange(desc(MISSING)) %>%
  select(VAR, TYPE, MISSING)

write_csv(vars_missing, paste(file.dir,"vars_missing.csv",sep = ""))

# Visual of missing values by each variable
ggplot(data=vars_missing[vars_missing$MISSING!=0,], aes(x=VAR, y=MISSING)) +
  geom_bar(stat="sum", aes(fill=VAR))  + coord_flip() + theme_minimal()+ theme(legend.position = "none") +
  labs(title="Missing Values in Data", x="Number Missing", y="Variable")

#Collinearity detection
fullmod <- glm(INS ~., data = insurance, family = binomial(link = "logit"))

vifs <- data.frame(vif(fullmod)[,1])%>%
  rename(VIF=vif.fullmod....1.)
vifs$VAR <- row.names(vifs)
vifs <- vifs %>% select(VAR, VIF) %>% filter(VIF > 5)

insur_sm <- insurance %>%
  select(ILS, ILSBAL, MM, MMBAL, MTG, MTGBAL, CC, CCBAL)

insur_sm %>%
  filter(ILS==0 & ILSBAL == 0) %>%
  nrow()

```

For people with installment loan (ISL), their ILSBAL is nonzero. Same with MM and MMBAL.




